---
title: "Exercises - Class 4: Functions"
author: "Sander Bossier"
date: "`r Sys.Date()`"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## 1.

Harmonic numbers are sums of the reciprocals of the positive integers. More specifically, the n’th harmonic number is:
$$\sum_{i=1}^{n} \frac{1}{i}$$
Create a function `Harmonic(n)` which computes the nth harmonic number. The body of this function can consist of a single line. An example is:
```{r}
Harmonic <- function(n) sum(1 / (1:n))

Harmonic(5)
```

## 2.

Harmonic numbers are recursive:
$$H_{n} = H_{n-1} + \frac{1}{n}$$
Write a recursive function rHarmonic(n) which computes the nth harmonic number recursively (in other words, the result should be the same as in the previous exercise).

```{r}
rHarmonic <- function(n) {
  if (n == 1) {
    1
  }
  else{
    rHarmonic(n - 1) + (1 / n)
  }
}
rHarmonic(5)
```

## 3.

Compare the computation time of `Harmonic(n)` and `rHarmonic(n)` for the 1000th harmonic number (using
R’s core timing function).

```{r}
system.time(
  Harmonic(1000)
)
system.time(
  rHarmonic(1000)
)
system.time(
  Harmonic(1e6)
)

```

## 4.

Create a binary infix operator __`%fill%`__ which replaces all __`NA`__ values in the vector on the left-hand side by the value specified on the right-hand side.

```{r}
`%fill%` <- function(vec, value) {
  vec[is.na(vec)] <- value # or
  #replace(vec, is.na(vec1), value)
  vec
}

vec1 <- c(5, 55, NA, -9, -99, NA, NA, -5, -9, 999)
vec2 <- vec1 %fill% 0
vec2
```

## 5.

Create a replacement function fill() which replaces all NA values in a vector by the value specified on the right-hand side.

```{r}
`fill<-` <- function(vec, value) {
  vec[is.na(vec)] <- value
  vec
}

vec1 <- c(5, 55, NA, -9, -99, NA, NA, -5, -9, 999)
fill(vec1) <- 0
vec1
```

## 6.

Compare the computation time of the %fill% operator and the fill() replacement function on the following large vector (you can copy-paste these two lines):

```{r}
library(microbenchmark)
set.seed(123)
vec3 <- sample(c(5, -9, NA), size = 1e6, replace = TRUE)
```

Do this using both R’s core function and the package microbenchmark (replace the NA’s by 0).

```{r}
system.time(
  tmp <- vec3 %fill% 0
)
system.time(
  fill(vec3) <- 0
)

microbenchmark(
  tmp <- vec3 %fill% 0,
  unit = "us"
)
microbenchmark(
  fill(vec3) <- 0,
  unit = "us"
)
```

## 7.

Create two functions __`full1()`__ and __`full2()`__ which both extract the non-NA elements from a vector. One of both functions needs to make use of the functional `Filter()` (see last week). Compare their computation times on the `vec3` object from the previous exercises. You can use both R’s core function and the __microbenchmark__ package but for the latter you can specify a number of `times` <= 100.

Note: There is even a way to combine the `Filter()` function with the `Negate()` function. If you manage to find that one, then you can compare all three functions.

```{r}
full1 <- function(vec) {
  vec[!is.na(vec)]
}

full2 <- function(vec) {
  Filter(function(x) !is.na(x), vec)
}

full3 <- function(vec) {
  Filter(Negate(function(x) is.na(x)), vec)
}
```

```{r}
vec1 <- c(5, 55, NA, -9, -99, NA, NA, -5, -9, 999)
microbenchmark(
  full1(vec3),
  full2(vec3),
  full3(vec3),
  times = 100,
  unit = "ns"
)
```

## 8.

One of the features of R’s syntax is that a function name before the opening bracket __`(`__ can actually be a variable. This allows uses such as the following:

```{r}
f <- sum
f(vec2)

f <- mean
f(vec2)
```

Create a function `twofun()` which has two formal  arguments:

* `x` which can be a numeric vector
* `fun` which can be any function computing some descriptive statistic

The function `twofun()` should apply the function specified as `fun` to every two consecutive elements in the vector specified as `x`. In other words, if `fun = sum` then `twofun()` will compute the sum of every two consecutive elements in `x` like we have seen last week (see the last exercise of the previous week). However, `twofun()` allows for more flexibility:

```{r}
twofun <- function(vec, fun = sum) {
  vapply(
    seq_len(length(vec) - 1),
    FUN = function(i) fun(vec[i:(i + 1)]),
    numeric(1)
  )
}
```

```{r}
twofun(vec1, fun = sum)
twofun(vec1, fun = mean)
twofun(vec1, fun = median)
twofun(vec1, fun = var)
twofun(vec1)
```

## 9.

R’s functions for descriptive statistics have an argument `na.rm` and `var()` even has an argument `use`. Rewrite the `twosum()` function so that it can handle all of these extra arguments without any conflict.

```{r}
twofun <- function(vec, fun = sum, ...) {
  
  vapply(
    X = seq_len(length(vec) - 1),
    FUN = function(i) fun(vec[i:(i + 1)], ...),
    FUN.VALUE = numeric(1)
  )
}
```


## 10.

User-defined functions in R have the type `closure` because they can be defined within another function (see the Wikipedia page on closures but a better description can be found in chapter 10 of Hadley Wickham’s _Advanced R_). Create a function `twofunfun()` which takes a statistical function as its single argument `fun` and which _returns another function_ which applies `fun` to the two consecutive elements of any vector (i.e. like in the previous two exercises).

```{r}
twofunfun <- function(fun = sum) {
  out <- function(vec, ...) {
    vapply(
      X = seq_len(length(vec) - 1),
      FUN = function(i) fun(vec[i:(i + 1)], ...),
      FUN.VALUE = numeric(1)
    )
  }
  out
}
```

```{r}
twomean <- twofunfun(fun = mean)
twomean(vec1)

twovar <- twofunfun(fun = var)
twovar(vec1)
```

Because closures such as `twomean()` and/or `twovar()` are created within `twofunfun()`, they have the (temporary) execution environment of the latter as their environment:

```{r}
environment(twomean)
environment(twovar)
```

Such execution environments contain the values of the formal arguments (see slide 26) and that is why these environments should contain the value for the object `fun`. Show that this is indeed the case and get those values.

```{r}
exists("fun", environment(twomean))
get("fun", environment(twomean))
exists("fun", environment(twovar))
get("fun", environment(twovar))
```

## 11.

Write an S3 method `freqs()` which counts how often every element appears in a vector. It should work for the S3 class `text` which is simply an extension of character vectors:

```{r}
freqs <- function(x, ...) {
  UseMethod("freqs")
}

freqs.text <- function(vec) {
  table(vec, useNA = "ifany")
}
```

```{r}
freqskp <- function(x, ...) UseMethod("freqs")
freqskp.text <- function(x) vapply(unique(x),
                                 function(j) sum(x == j),
                                 numeric(1))
```


```{r}
vec4 <- c("a", "rose", "is", "a", "rose", "is", "a", "rose")
class(vec4) <- "text"
freqs(vec4)
```
```{r}
freqskp(vec4)
```

```{r}
microbenchmark(freqs(vec4),
               freqskp(vec4),
               unit = "ns")
```

## 12.

Write a similar method `count()` for the S4 class `txt4` as well one as for the RC class `txtR` (we need to use different class and method names because otherwise R will mix them up; it is standard practice to choose just one OOP system and stick to it throughout your project).

```{r}
library(methods)

setClass("txt4", contains = "character")

setGeneric("count", function(vec) standardGeneric("count"))

setMethod("count", "txt4", function(vec) table(vec, useNA = "ifany"))
```

```{r}
vec5 <- new("txt4", c("a", "rose", "is", "a", "rose", "is", "a", "rose"))
count(vec5)
```

```{r}
setRefClass(
  "txtR",
  fields = c(vec = "character"),
  methods = list(
    count = function() {
      table(vec, useNA = "ifany")
    }
  )
  )
```

```{r}
vec6 <- new("txtR", vec = c("a", "rose", "is", "a", "rose", "is", "a", "rose"))
vec6$count()
```

## 13.

Compare the computation time of the three methods with the computation time of simply using the `table()` function (which you might already have used in your methods but that is okay: in that case you will actually measure the computation time of the method dispatch). Do this comparison on a vector of a million randomly sampled letters from the (Roman) alphabet:

```{r}
set.seed(123)

vec7 <- sample(letters, 1e6, replace = TRUE)

vec8 <- new("txt4", vec7)
vec9 <- new("txtR", vec = vec7)
class(vec7) <- "text"
```

```{r}
microbenchmark(
  freqs(vec7),
  count(vec8),
  vec9$count()
)
```

## 14.

A simple linear regression model estimates an __intercept__ and a __slope__ parameter of a regression line with which one variable can be explained/predicted in terms of another variable. For instance, the core R dataset `cars` has two variables _speed (mph)_ and _stopping distance (ft)_ and the second variable can be predicted/explained in terms of the first variable using the following intercept and slope parameters (estimated on the basis of 50 observations):

```{r}
linmod <- lm(dist ~ speed, data = cars)
linmod
```

In other words, this linear regression model essentially states (using more digits for both parameters):

$$dist = `r coef(linmod)[1]` + `r coef(linmod)[2]` \, \times \, speed$$

This means that particular __predictions__ for _stopping distance (ft)_ can be obtained by specifying certain values for _speed (mph)_ and plugging them into the equation of the linear regression model. Say, we would like the predicted values of _stopping distance (ft)_ for the _speed (mph)_ values _5_, _6_ and _7_. Modify the following two functions in order to obtain the proper result instead of the error message:

```{r}
parameters <- function(mod) {
coefs <- coefficients(mod)
}
predictions <- function(x, mod) {
parameters(mod)
coefs[1] + coefs[2] * x
}
predictions(x = 5:7, mod = linmod)
```

```{r}
parameters <- function(mod) {
  coefs <- coefficients(mod)
}

predictions <- function(x, mod) {
  coefs <- parameters(mod)
  coefs[1] + coefs[2] * x
}

predictions(x = 5:7, mod = linmod)
```