---
title: "Exercises - Class 4: Functions"
author: "Sander Bossier"
date: "`r Sys.Date()`"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## 1.

Harmonic numbers are sums of the reciprocals of the positive integers. More specifically, the n’th harmonic number is:
$$\sum_{i=1}^{n} \frac{1}{i}$$
Create a function `Harmonic(n)` which computes the nth harmonic number. The body of this function can consist of a single line. An example is:
```{r}
Harmonic <- function(n) sum(1 / (1:n))

Harmonic(5)
```

## 2.

Harmonic numbers are recursive:
$$H_{n} = H_{n-1} + \frac{1}{n}$$
Write a recursive function rHarmonic(n) which computes the nth harmonic number recursively (in other words, the result should be the same as in the previous exercise).

```{r}
rHarmonic <- function(n) {
  if (n == 1) {
    1
  }
  else{
    rHarmonic(n - 1) + (1 / n)
  }
}
rHarmonic(5)
```

## 3.

Compare the computation time of `Harmonic(n)` and `rHarmonic(n)` for the 1000th harmonic number (using
R’s core timing function).

```{r}
system.time(
  Harmonic(1000)
)
system.time(
  rHarmonic(1000)
)
system.time(
  Harmonic(1e6)
)

```

## 4.

Create a binary infix operator __`%fill%`__ which replaces all __`NA`__ values in the vector on the left-hand side by the value specified on the right-hand side.

```{r}
`%fill%` <- function(vec, value) {
  vec[is.na(vec)] <- value # or
  #replace(vec, is.na(vec1), value)
  vec
}

vec1 <- c(5, 55, NA, -9, -99, NA, NA, -5, -9, 999)
vec2 <- vec1 %fill% 0
vec2
```

## 5.

Create a replacement function fill() which replaces all NA values in a vector by the value specified on the right-hand side.

```{r}
`fill<-` <- function(vec, value) {
  vec[is.na(vec)] <- value
  vec
}

vec1 <- c(5, 55, NA, -9, -99, NA, NA, -5, -9, 999)
fill(vec1) <- 0
vec1
```

## 6.

Compare the computation time of the %fill% operator and the fill() replacement function on the following large vector (you can copy-paste these two lines):

```{r}
library(microbenchmark)
set.seed(123)
vec3 <- sample(c(5, -9, NA), size = 1e6, replace = TRUE)
```

Do this using both R’s core function and the package microbenchmark (replace the NA’s by 0).

```{r}
system.time(
  tmp <- vec3 %fill% 0
)
system.time(
  fill(vec3) <- 0
)

microbenchmark(
  tmp <- vec3 %fill% 0,
  unit = "us"
)
microbenchmark(
  fill(vec3) <- 0,
  unit = "us"
)
```

## 7.

Create two functions __`full1()`__ and __`full2()`__ which both extract the non-NA elements from a vector. One of both functions needs to make use of the functional `Filter()` (see last week). Compare their computation times on the `vec3` object from the previous exercises. You can use both R’s core function and the __microbenchmark__ package but for the latter you can specify a number of `times` <= 100.

Note: There is even a way to combine the `Filter()` function with the `Negate()` function. If you manage to find that one, then you can compare all three functions.

```{r}
full1 <- function(vec) {
  vec[!is.na(vec)]
}

full2 <- function(vec) {
  Filter(function(x) !is.na(x), vec)
}

full3 <- function(vec) {
  Filter(Negate(function(x) is.na(x)), vec)
}
```

```{r}
vec1 <- c(5, 55, NA, -9, -99, NA, NA, -5, -9, 999)
microbenchmark(
  full1(vec3),
  full2(vec3),
  full3(vec3),
  times = 100,
  unit = "ns"
)
```

## 8.

One of the features of R’s syntax is that a function name before the opening bracket __`(`__ can actually be a variable. This allows uses such as the following:

```{r}
f <- sum
f(vec2)

f <- mean
f(vec2)
```

Create a function `twofun()` which has two formal  arguments:

* `x` which can be a numeric vector
* `fun` which can be any function computing some descriptive statistic

The function `twofun()` should apply the function specified as `fun` to every two consecutive elements in the vector specified as `x`. In other words, if `fun = sum` then `twofun()` will compute the sum of every two consecutive elements in `x` like we have seen last week (see the last exercise of the previous week). However, `twofun()` allows for more flexibility:

```{r}
twofun <- function(vec, fun = sum) {
  vapply(
    seq_len(length(vec) - 1),
    FUN = function(i) fun(vec[i:(i + 1)]),
    numeric(1)
  )
}
```

```{r}
twofun(vec1, fun = sum)
twofun(vec1, fun = mean)
twofun(vec1, fun = median)
twofun(vec1, fun = var)
twofun(vec1)
```

## 9.

R’s functions for descriptive statistics have an argument `na.rm` and `var()` even has an argument `use`. Rewrite the `twosum()` function so that it can handle all of these extra arguments without any conflict.

```{r}
twofun <- function(vec, fun = sum, ...) {
  
  vapply(
    X = seq_len(length(vec) - 1),
    FUN = function(i) fun(vec[i:(i + 1)], ...),
    FUN.VALUE = numeric(1)
  )
}
```


## 10.



```{r}
twofunfun <- function(fun = sum) {
  out <- function(vec, ...) {
    vapply(
      X = seq_len(length(vec) - 1),
      FUN = function(i) fun(vec[i:(i + 1)], ...),
      FUN.VALUE = numeric(1)
    )
  }
  out
}
```

```{r}
twomean <- twofunfun(fun = mean)
twomean(vec1)

twovar <- twofunfun(fun = var)
twovar(vec1)
```

```{r}
environment(twomean)
environment(twovar)
```

```{r}
exists("fun", environment(twomean))
get("fun", environment(twomean))
exists("fun", environment(twovar))
get("fun", environment(twovar))
```

