---
title: "sb_rmd_ex_05"
author: "Sander Bossier"
date: "`r Sys.Date()`"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(microbenchmark)
library(tidyverse)
library(data.table)
```

## 1.

```{r results='hide'}
speed_read_baseR <- microbenchmark(
  read.delim(
    file = "https://raw.githubusercontent.com/kplevoet/texts/main/doyle/baskervilles_TAG.txt",
    sep = "\t",
    encoding = "UTF-8",
    quote = ""
    ),
  times = 20,
  unit = "ms"
)

speed_read_tidy <- microbenchmark(
  read_delim(
    file = "https://raw.githubusercontent.com/kplevoet/texts/main/doyle/baskervilles_TAG.txt",
    show_col_types = FALSE,
    quote = ""
  ),
  times = 20,
  unit = "ms"
)

speed_read_dt <- microbenchmark(
  fread(
    input = "https://raw.githubusercontent.com/kplevoet/texts/main/doyle/baskervilles_TAG.txt",
    sep = "\t",
    encoding = "UTF-8",
    quote = ""
  ),
  times = 20,
  unit = "ms"
)
```

The average speed of the base R function `read.delim` is `r summary(speed_read_baseR)$mean` ms, the average speed of the tidyverse readr `read_delim` function is `r summary(speed_read_tidy)$mean` ms, and the average speed of the data.table `fread` function is `r summary(speed_read_dt)$mean` ms.

```{r}
data_baseR <- read.delim(
  file = "https://raw.githubusercontent.com/kplevoet/texts/main/doyle/baskervilles_TAG.txt",
  sep = "\t",
  encoding = "UTF-8",
  quote = ""
  )

data_tidy <- read_delim(
  file = "https://raw.githubusercontent.com/kplevoet/texts/main/doyle/baskervilles_TAG.txt",
  show_col_types = FALSE,
  quote = ""
  )

data_dt <- fread(
  input = "https://raw.githubusercontent.com/kplevoet/texts/main/doyle/baskervilles_TAG.txt",
  sep = "\t",
  encoding = "UTF-8",
  quote = ""
)
```

The object size for the base R data.frame is `r format(object.size(data_baseR), units = "Kb")`, the object size of of tidyverse tibble is `r format(object.size(data_tidy), units = "Kb")`, and the size of the data.frame is `r format(object.size(data_dt), units = "Kb")`.

## 2.

```{r}
data_baseR_noPUNCT <- subset(
  x = data_baseR,
  subset = toupper(upos) != "PUNCT"
)

data_tidy_noPUNCT <- data_tidy %>%
  filter(toupper(upos) != "PUNCT")

data_dt_noPUNCT <- data_dt[toupper(upos) != "PUNCT"]
```

## 3.

```{r}
data_baseR_noPUNCT <- within(
  data = data_baseR_noPUNCT,
  expr = {
    token_low <- tolower(token)
    lemma_low <- tolower(lemma)
  }
)

data_tidy_noPUNCT <- data_tidy_noPUNCT %>%
  mutate(
    token_low = tolower(token),
    lemma_low = tolower(lemma)
  )

data_dt_noPUNCT[
  ,
  c("token_low", "lemma_low") := list(
    tolower(token),
    tolower(lemma)
  )
]
```


## 4.

```{r}
lemma_low_freq_baseR <- with(
    data = data_baseR_noPUNCT,
    expr = as.data.frame(
      x = table(
        lemma_low
        ),
      responseName = "Freq"
    )
  )

lemma_low_freq_tidy <- data_tidy_noPUNCT %>%
  group_by(lemma_low) %>%
  summarise(Freq = n())

lemma_low_freq_dt <- data_dt_noPUNCT[
  ,
  list(Freq = .N),
  by = lemma_low
]
```

## 5.

```{r}
lemma_low_freq_baseR <- lemma_low_freq_baseR[
  with(
    data = lemma_low_freq_baseR,
    expr = order(Freq,
                 decreasing = TRUE
                 )
    ),
  ]

lemma_low_freq_tidy <- lemma_low_freq_tidy %>%
  arrange(desc(Freq))

lemma_low_freq_dt <- lemma_low_freq_dt[
  order(Freq, decreasing = TRUE)
]
```

## 6.

```{r}
upos_type <- data.frame(upos = c("NOUN", "PROPN", "ADJ",
                                 "VERB", "ADV", "NUM",
                                 "DET", "PRON", "ADP",
                                 "CCONJ", "SCONJ", "PART",
                                 "AUX", "INTJ"),
                        type = c(rep("Content", 6),
                                 rep("Function", 8)))


data_baseR_noPUNCT$upos <- toupper(data_baseR_noPUNCT$upos)
data_baseR_noPUNCT_merge <- merge(
  x = data_baseR_noPUNCT,
  y = upos_type,
  by.x = "upos",
  by.y = "upos"
  )

data_tidy_noPUNCT_merge <- data_tidy_noPUNCT %>%
  mutate(upos = str_to_upper(upos)) %>%
  inner_join(
    x = .,
    y = upos_type,
    by = "upos"
  )

data_dt_noPUNCT <- data_dt_noPUNCT[as.data.table(upos_type),
                on = c(upos = "upos")]
```

## 7.

```{r}
lemma_low_freq_baseR$Perc <- lemma_low_freq_baseR$Freq / sum(lemma_low_freq_baseR$Freq) * 100

lemma_low_freq_baseR_long <- reshape(
    lemma_low_freq_baseR,
    direction = "long",
    varying = c("Freq", "Perc"),
    idvar = "lemma_low",
    v.names = "Value",
    times = c("Freq", "Perc"),
    timevar = "Type"
  )

lemma_low_freq_tidy_long <- lemma_low_freq_tidy %>%
  mutate(
    Perc = Freq / sum(Freq) * 100
  ) %>%
  pivot_longer(
    cols = c("Freq", "Perc"),
    names_to = "Type",
    values_to = "Value"
  )

lemma_low_freq_dt <- lemma_low_freq_dt[
    ,
    Perc := Freq / sum(Freq) * 100
  ]

lemma_low_freq_dt_long <- melt(
    data = lemma_low_freq_dt,
    id.vars = "lemma_low",
    measure.vars = c("Freq", "Perc"),
    variable.name = "Type",
    value.name = "Value"
  )
```
