---
title: "sb_rmd_ex_07"
author: "Sander Bossier"
date: "`r Sys.Date()`"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## 1.

Consider the character vector x:

```{r}
x <- c("Swan swam over the pond, Swim swan swim!",
"Swam swam back again - Well swum swan!")
```

Split the strings in `x` into words, removing any *commas* or *hyphens*. Call the result `y`.

```{r}
y <- strsplit(
  x,
  split = ",? -? ?"
)
y
```
## 2.

The .rda file `hafu` in the RStudio Project `Text` contains a data set `hafu` on manga characters. In the `Father` and `Mother` columns, some values have question marks after the country name, indicating that the author was uncertain about the nationality of the parent:

```{r}
load("hafu.rda")
head(hafu[, -ncol(hafu)], n = 10)
```
Create two vectors `FatherUncertain` and `MotherUncertain` denoting whether or not there was a question mark in the `Father` or `Mother` column, respectively.

```{r}
FatherUncertainSB <- endsWith(
  x = as.character(hafu$Father),
  suffix = "?"
)
MotherUncertainSB <- endsWith(
  x = as.character(hafu$Mother),
  suffix = "?"
)

FatherUncertain <- with(hafu, grepl("\\?", Father))
MotherUncertain <- with(hafu, grepl("\\?", Mother))

identical(FatherUncertainSB, FatherUncertain)
identical(MotherUncertainSB, MotherUncertain)
```

## 3.

Create a new data frame `hafu2` by removing the question marks from the `Father` and `Mother` columns in `hafu` (i.e. replace them by an empty string).

```{r}
hafu2 <- within(hafu, {
  Father <- gsub(pattern = "\\?", replacement = "", x = Father)
  Mother <- gsub(pattern = "\\?", replacement = "", x = Mother)
  }
)
```

## Challenges

## 4.

The .rda file `Dataline` contains a vector `Dataline` that contains a single value. This value is the result of a
webscraper. It contains 1500 fields, separated by a semicolon:

```{r}
load("Dataline.rda")
```

Create a vector `Fields` which contains the 1500 fields as separate values.

```{r}
Fields <- strsplit(Dataline, split = ";")[[1]]
```

## 5.

On the basis of the vector `Fields` create:
* A character vector `Label` containing the labels of every field: possible values are “Low”, “Middle” and “High”.
* A numeric vector `ID` containing the ID numbers of every field (i.e. the value following the underscore).
* A character vector `newLabel` containing labels constructed as `ID_xxx_Label`, with `xxx` the value of `ID` for that field and `Label` the value of `Label` for that field.

```{r}
head(Fields, n = 10)
```

```{r}
Label <- gsub(
  pattern = "ID_\\d+", #KP solution is more general by no requiring the underscore followed by any number of digits, but by just specifying any number of characters (zero or more occurrences)
  replacement = "",
  x = Fields
)

ID <- as.numeric(
  gsub(
    pattern = "\\w+ID_", # again KP solution is more general as it matches a sequence of one or more non-underscore characters, followed by an underscore
    replacement = "",
    x = Fields
  )
)

newLabel <- paste(
  "ID",
  ID,
  Label,
  sep = "_"
)
```

```{r Fields}
# Solution 5:
LabelKP <- gsub("ID.*", "", Fields)
IDKP <- as.numeric(gsub("[^_]+_", "", Fields))
newLabelKP <- paste("ID", ID, Label, sep = "_")
```

```{r}
identical(Label, LabelKP)
identical(ID, IDKP)
identical(newLabel, newLabelKP)
```

## 6.

The file genes.rda contains a named vector `genes`. This vector contains the function descriptions of a set of genes. The names of this vector give the COG codes of these genes:

```{r}
load("genes.rda")
head(genes, n = 10)
```

Use this vector to create a character vector `geneclass` with the values “Predicted”, “Uncharacterized” and “Described”. These are defined as follows:
* If the gene description contains “Uncharacterized”, it is always called “Uncharacterized”
* If the gene description contains “Predicted” but not “Uncharacterized”, it is called “Predicted”
* All other genes are called “Described”

```{r}
geneclass <- rep("Described", length(genes))
IDunc <- grep(pattern = "uncharacterized", x = genes, ignore.case = TRUE)
IDpred <- grep(pattern = "predicted", x = genes, ignore.case = TRUE)
geneclass[IDpred] <- "Predicted"
geneclass[IDunc] <- "Uncharacterized"
```

## 7.

Use the `genes` vector to find all COG codes for functions containing an enzyme name. An enzyme can be recognised by their name ending on “ase” or “ases”. For this, you need to:
* Extract the COG code from the names: this consists of the numbers following the letters COG in the
name
* Find at which positions in the vector genes an enzyme is described
* Use the information from the previous two steps to find only the COG codes of the enzymes

```{r}
COGcodes <- gsub(
  pattern = ".*COG",
  replacement = "",
  x = names(genes)
)

```

