---
title: "Exercises - Class 2; Objects and classes"
author: "Sander Bossier"
date: "2023-10-10"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## 1.

Create the following vectors with the function `rep()` (or its variants described on its help page):

```{r}
rep(3, times = 5)

rep(c(3, 1, 2), each = 2)

rep(c(3, 1, 2), length.out = 5)

rep(c(3, 1, 2), times = c(3, 1, 2))
```

## 2.

Create the following vectors with the function `seq()` (or its variants described on its help page):

```{r}
seq(from = 1, to = 9, by = 2)

seq(from = 1, to = 10, along.with = c(3, 2, 1))

seq(from = 1, to = 10, length.out = 7)

seq(from = 10, to = 2, by = -2)
```

## 3.

The function `replicate()` is sometimes wrongly used instead of `rep()` to repeat values. Use the function
`system.time()` to measure the performance of both functions in repeating the value 3 for a total of 1e6
times.

```{r}
system.time(
  rep(1, times = 1e6)
)

system.time(
  replicate(1e6, 1)
)
```

## 4.

The difference between `integer` and `double`(-precision floating point number) values is the byte size with which R stores them in memory. Byte size is also dependent on the attributes of the object. Use the function object.size() to show this for:
List these six objects in increasing order of byte size in your R code.

```{r}
obj01 <- 1:10
obj02 <- seq(1, 10, by = 1)
obj03 <- matrix(1:10, nrow = 5)
obj04 <- matrix(seq(1, 10, by = 1), nrow = 5)
obj05 <- 1:10
class(obj05) <- "humpty"
obj06 <- seq(1, 10, by = 1)
class(obj06) <- "dumpty"


print(c("obj01 size:", object.size(obj01)))
print(c("obj02 size:", object.size(obj02)))
print(c("obj03 size:", object.size(obj03)))
print(c("obj05 size:", object.size(obj05), "class:", class(obj05)))
print(c("obj04 size:", object.size(obj04)))
print(c("obj06 size:", object.size(obj06), "class:", class(obj06)))
```

## 5.

Rank the following objects/vectors according to the byte size. Which result surprises you?

```{r}
object.size(rep(c(TRUE, FALSE), 5))
object.size(seq(1, 10, by = 1))
object.size(vector(mode = "logical", length = 10))
object.size(vector(mode = "numeric", length = 10))
object.size(vector(mode = "list", length = 10))
object.size(vector(mode = "character", length = 10))
```

## 6.

```{r}
object.size(NULL)
object.size(NA)
obj07 <- c(label = NULL)
obj08 <- c(label = NA)
object.size(obj07)
object.size(obj08)
```

## 7.

```{r}
as.numeric(2 * log(1i) / 1i) == pi
```

## 8.

```{r}
alphabetS4 <- setClass("alphabetS4", slots = c(symbols = "character", size = "numeric", type = "character"))

obj09 <- new("alphabetS4", symbols = letters, size = length(letters), type = "roman")

obj09

alphabetRC <- setRefClass("alphabetRC", fields = c(symbols = "character", size = "numeric", type = "character"))

obj10 <- new("alphabetRC", symbols = letters, size = length(letters), type = "roman")

obj10
```

## 9.

```{r}
obj11 <- list(
  symbols = letters,
  size = length(letters),
  type = "roman"
)

obj12 <- new.env()
obj12$symbols <- letters
obj12$size <- length(letters)
obj12$type <- "roman"
```

```{r}
object.size(obj09)
object.size(obj10)
object.size(obj11)
object.size(obj12)
```

## 10.

```{r}
vecX <- c(symbols = "a", size = "1", type = "roman")
expY <- expression(vecY <- c(symbols = "a", size = "1", type = "roman"))
eval(expY)
vecX == vecY
identical(vecX, vecY)
object.size(vecX)
object.size(vecY)
```

## 11.

```{r}
err <- simpleError("Watch out for this!")
object.size(err)
object.size("Watch out for this!")
attributes(err)
```
##12.

```{r}
is.atomic(err)
is.recursive(err)
```

