---
title: "Exercises - Class 2; Objects and classes"
author: "Sander Bossier"
date: "2023-10-10"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## 1.

Create the following vectors with the function `rep()` (or its variants described on its help page):

```{r}
rep(3, times = 5)

rep(c(3, 1, 2), each = 2)

rep(c(3, 1, 2), length.out = 5)

rep(c(3, 1, 2), times = c(3, 1, 2))
```

## 2.

Create the following vectors with the function `seq()` (or its variants described on its help page):

```{r}
seq(from = 1, to = 9, by = 2)

seq(from = 1, to = 10, along.with = c(3, 2, 1))

seq(from = 1, to = 10, length.out = 7)

seq(from = 10, to = 2, by = -2)
```

## 3.

The function `replicate()` is sometimes wrongly used instead of `rep()` to repeat values. Use the function
`system.time()` to measure the performance of both functions in repeating the value 3 for a total of 1e6
times.

```{r}
system.time(
  rep(1, times = 1e6)
)

system.time(
  replicate(1e6, 1)
)
```

## 4.

The difference between `integer` and `double`(-precision floating point number) values is the byte size with which R stores them in memory. Byte size is also dependent on the attributes of the object. Use the function object.size() to show this for:
List these six objects in increasing order of byte size in your R code.

```{r}
obj01 <- 1:10
obj02 <- seq(1, 10, by = 1)
obj03 <- matrix(1:10, nrow = 5)
obj04 <- matrix(seq(1, 10, by = 1), nrow = 5)
obj05 <- as.character(1:10)
obj06 <- as.raw(seq(1, 10, by = 1))

print(c("obj06", object.size(obj06)))
print(c("obj01", object.size(obj01)))
print(c("obj02", object.size(obj02)))
print(c("obj03", object.size(obj03)))
print(c("obj04", object.size(obj04)))
print(c("obj05", object.size(obj05)))
```

## 5.
