---
title: "Edgar Anderson’s Iris Data"
output:
  pdf_document:
header-includes:
  - "\\makeatletter"
  - "\\renewcommand*\\l@section[2]{\\@dottedtocline{1}{1.5em}{2.3em}{#1}{#2}\\addvspace{0.7em}}"
  - "\\makeatother"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

\tableofcontents

\section{Description}
This famous (Fisher’s or Anderson’s) iris data set gives the measurements in centimeters of the variables
sepal length and width and petal length and width, respectively, for 50 flowers from each of 3 species of iris.
The species are _Iris setosa_, _versicolor_, and _virginica_.


\section{Usage}
`iris`

\section{Format}
`iris` is a data frame with 150 cases (rows) and 5 variables (columns) named:

- `Sepal.Length`
- `Sepal.Width`
- `Petal.Length`
- `Petal.Width`
- `Species`

\section{Source}
Anderson, Edgar (1935). "The irises of the Gaspe Peninsula.", _Bulletin of the American Iris Society_, __59__: 2–5.
Fisher, Ronald A. (1936). "The use of multiple measurements in taxonomic problems.", _Annals of Eugenics_, __7__ (Part II): 179–188.

\section{Examples}
We investigate the Sepal and Petal leaves for the three species in the Iris data:

```{r}
summary(iris)
```

To examine the Sepal leaves, we select the length and the width:

```{r results='hide'}
llen <- iris$Sepal.Length
lwid <- iris$Sepal.Width
```

Then we plot the data:

```{r}
plot(llen, lwid, xlab = "Length", ylab = "Width",
     pch = 20, col = as.numeric(iris$Species))
legend("bottomright", legend = levels(iris$Species), col = 1:3, pch = 20)
```

We can also select the Petal leaves:

```{r results='hide'}
llen <- iris$Petal.Length
lwid <- iris$Petal.Width
```

This gives us the following plot:

```{r echo=FALSE}
plot(llen, lwid, xlab = "Length", ylab = "Width",
     pch = 20, col = as.numeric(iris$Species))
legend("bottomright", legend = levels(iris$Species), col = 1:3, pch = 20)
```