---
title: "sb_rmd_ex_07"
author: "Sander Bossier"
date: "`r Sys.Date()`"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## 1.

Consider the character vector x:

```{r}
x <- c("Swan swam over the pond, Swim swan swim!",
"Swam swam back again - Well swum swan!")
```

Split the strings in `x` into words, removing any *commas* or *hyphens*. Call the result `y`.

```{r}
y <- strsplit(
  x,
  split = ",? -? ?"
)
y
```
## 2.

The .rda file `hafu` in the RStudio Project `Text` contains a data set `hafu` on manga characters. In the `Father` and `Mother` columns, some values have question marks after the country name, indicating that the author was uncertain about the nationality of the parent:

```{r}
load("hafu.rda")
head(hafu[, -ncol(hafu)], n = 10)
```
Create two vectors `FatherUncertain` and `MotherUncertain` denoting whether or not there was a question mark in the `Father` or `Mother` column, respectively.

```{r}
FatherUncertainSB <- endsWith(
  x = as.character(hafu$Father),
  suffix = "?"
)
MotherUncertainSB <- endsWith(
  x = as.character(hafu$Mother),
  suffix = "?"
)

FatherUncertain <- with(hafu, grepl("\\?", Father))
MotherUncertain <- with(hafu, grepl("\\?", Mother))

identical(FatherUncertainSB, FatherUncertain)
identical(MotherUncertainSB, MotherUncertain)
```

## 3.

Create a new data frame `hafu2` by removing the question marks from the `Father` and `Mother` columns in `hafu` (i.e. replace them by an empty string).

```{r}
hafu2 <- within(hafu, {
  Father <- gsub(pattern = "\\?", replacement = "", x = Father)
  Mother <- gsub(pattern = "\\?", replacement = "", x = Mother)
  }
)
```

## Challenges

## 4.

The .rda file `Dataline` contains a vector `Dataline` that contains a single value. This value is the result of a
webscraper. It contains 1500 fields, separated by a semicolon:

```{r}
load("Dataline.rda")
```

Create a vector `Fields` which contains the 1500 fields as separate values.

```{r}
Fields <- strsplit(Dataline, split = ";")[[1]]
```

## 5.

On the basis of the vector `Fields` create:
* A character vector `Label` containing the labels of every field: possible values are “Low”, “Middle” and “High”.
* A numeric vector `ID` containing the ID numbers of every field (i.e. the value following the underscore).
* A character vector `newLabel` containing labels constructed as `ID_xxx_Label`, with `xxx` the value of `ID` for that field and `Label` the value of `Label` for that field.

```{r}
head(Fields, n = 10)
```

```{r}

```

