---
title: "sb_rmd_ex_07"
author: "Sander Bossier"
date: "`r Sys.Date()`"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## 1.

Consider the character vector x:

```{r}
x <- c(
  "Swan swam over the pond, Swim swan swim!",
  "Swam swam back again - Well swum swan!"
)
```

Split the strings in `x` into words, removing any *commas* or *hyphens*. Call the result `y`.

```{r}
y <- strsplit(
  x,
  split = ",? -? ?"
)
y
```
## 2.

The .rda file `hafu` in the RStudio Project `Text` contains a data set `hafu` on manga characters. In the `Father` and `Mother` columns, some values have question marks after the country name, indicating that the author was uncertain about the nationality of the parent:

```{r}
load("hafu.rda")
head(hafu[, -ncol(hafu)], n = 10)
```
Create two vectors `FatherUncertain` and `MotherUncertain` denoting whether or not there was a question mark in the `Father` or `Mother` column, respectively.

```{r}
FatherUncertainSB <- endsWith(
  x = as.character(hafu$Father),
  suffix = "?"
)
MotherUncertainSB <- endsWith(
  x = as.character(hafu$Mother),
  suffix = "?"
)

FatherUncertain <- with(hafu, grepl("\\?", Father))
MotherUncertain <- with(hafu, grepl("\\?", Mother))

identical(FatherUncertainSB, FatherUncertain)
identical(MotherUncertainSB, MotherUncertain)
```

## 3.

Create a new data frame `hafu2` by removing the question marks from the `Father` and `Mother` columns in `hafu` (i.e. replace them by an empty string).

```{r}
hafu2 <- within(hafu, {
  Father <- gsub(pattern = "\\?", replacement = "", x = Father)
  Mother <- gsub(pattern = "\\?", replacement = "", x = Mother)
  }
)
```

## Challenges

## 4.

The .rda file `Dataline` contains a vector `Dataline` that contains a single value. This value is the result of a
webscraper. It contains 1500 fields, separated by a semicolon:

```{r}
load("Dataline.rda")
```

Create a vector `Fields` which contains the 1500 fields as separate values.

```{r}
Fields <- strsplit(Dataline, split = ";")[[1]]
```

## 5.

On the basis of the vector `Fields` create:
* A character vector `Label` containing the labels of every field: possible values are “Low”, “Middle” and “High”.
* A numeric vector `ID` containing the ID numbers of every field (i.e. the value following the underscore).
* A character vector `newLabel` containing labels constructed as `ID_xxx_Label`, with `xxx` the value of `ID` for that field and `Label` the value of `Label` for that field.

```{r}
head(Fields, n = 10)
```

```{r}
Label <- gsub(
  pattern = "ID_\\d+", #KP solution is more general by no requiring the underscore followed by any number of digits, but by just specifying any number of characters (zero or more occurrences)
  replacement = "",
  x = Fields
)

ID <- as.numeric(
  gsub(
    pattern = "\\w+ID_", # again KP solution is more general as it matches a sequence of one or more non-underscore characters, followed by an underscore
    replacement = "",
    x = Fields
  )
)

newLabel <- paste(
  "ID",
  ID,
  Label,
  sep = "_"
)
```

```{r Fields}
# Solution 5:
LabelKP <- gsub("ID.*", "", Fields)
IDKP <- as.numeric(gsub("[^_]+_", "", Fields))
newLabelKP <- paste("ID", ID, Label, sep = "_")
```

```{r}
identical(Label, LabelKP)
identical(ID, IDKP)
identical(newLabel, newLabelKP)
```

## 6.

The file genes.rda contains a named vector `genes`. This vector contains the function descriptions of a set of genes. The names of this vector give the COG codes of these genes:

```{r}
load("genes.rda")
head(genes, n = 10)
```

Use this vector to create a character vector `geneclass` with the values “Predicted”, “Uncharacterized” and “Described”. These are defined as follows:
* If the gene description contains “Uncharacterized”, it is always called “Uncharacterized”
* If the gene description contains “Predicted” but not “Uncharacterized”, it is called “Predicted”
* All other genes are called “Described”

```{r}
geneclass <- rep("Described", length(genes))
IDunc <- grep(pattern = "uncharacterized", x = genes, ignore.case = TRUE)
IDpred <- grep(pattern = "predicted", x = genes, ignore.case = TRUE)
geneclass[IDpred] <- "Predicted"
geneclass[IDunc] <- "Uncharacterized"
```

## 7.

Use the `genes` vector to find all COG codes for functions containing an enzyme name. An enzyme can be recognised by their name ending on “ase” or “ases”. For this, you need to:
* Extract the COG code from the names: this consists of the numbers following the letters COG in the
name
* Find at which positions in the vector genes an enzyme is described
* Use the information from the previous two steps to find only the COG codes of the enzymes

```{r}
COGcodes <- gsub(
  pattern = ".*COG", # Again KP solutions is likely more general as it matches any non-digit characters
  replacement = "",
  x = names(genes)
)
postions <- grepl(
  pattern = ".*ases?\\b", # The questionmark only affects the 's' and thus the pattern can match 'ase' and 'ases'
  x = genes
)
COGcodes_recognized <- COGcodes[postions]
```

```{r}
COG <- gsub("\\D", "", names(genes))
isEnzyme <- grep("ases?\\b", genes)
enzymCOG <- COG[isEnzyme]
```

```{r}
identical(COGcodes_recognized, enzymCOG)
```

## 8.

The RStudio Project `Text` also contains two external files appletweets.txt and microsofttweets.txt. Each contains a selection of tweets scraped from Twitter/X and addressed to the helpdesk of either Apple or Microsoft.
Use the function `readLines()` to read in both files. Store the tweets for Apple in a vector `AppleTweets` and those for Microsoft in a vector `MSTweets.`

```{r}
AppleTweets <- readLines(
  con = "appletweets.txt",
  encoding = "UTF-8"
)
MSTweets <- readLines(
  con = "microsofttweets.txt",
  encoding = "UTF-8"
)
```

## 9.

Make sure that there are no empty character values ("") in either vector.

```{r}
AppleTweets <- readLines(
  con = "appletweets.txt",
  encoding = "UTF-8"
)
MSTweets <- readLines(
  con = "microsofttweets.txt",
  encoding = "UTF-8"
)
```

```{r}
AppleTweets <- AppleTweets[AppleTweets != ""]
MSTweets <- MSTweets[MSTweets != ""]
```

## 10.

Count the number of tweets that talk about iOS10. Think about possible spaces and capitalization.

```{r}
HASiOS10 <- sum(
  grepl(
    pattern = "\\bios\\s?10",
    x = AppleTweets,
    ignore.case = TRUE
  )
)
HASiOS10
```

# Extra exercises on dates and date-time objects

## 11.

Dates and Date-time objects in R are based on strings. Create the following objects (think about which type to use):
* A vector called time1: 2011-01-30
* A vector called time2: 30/01/11
* A vector called time3: Jan 30, 2011 14:45 UTC

You can you these strings as your starting point:

```{r}
text1 <- "2011-01-30"
text2 <- "30/01/11"
text3 <- "Jan 30, 2011 14:45 UTC"
```

```{r}
time1 <- as.Date(text1)
time2 <- strptime(text2, format = "%d/%m/%y")
time3 <- strptime(text3, format = "%b %d, %Y %H:%M", tz = "UTC")
time1
time2
time3
```

## 12.

Create a vector called `dateSeq` containing the dates:
* Starting from July, 1st 1989
* Ending at the last valid date before today
* With a periodicity of 3 months

```{r}
dateSeq <- seq(
  from = as.Date("July, 1st 1989", format = "%B, %dst %Y"),
  to = Sys.Date()-1,
  by = "3 months"
)
```

## 13.

Find in `dateSeq`:
* All dates that are later than January 2nd, 2000
* All dates in dateSeq that fall on a Monday

```{r}
dateSeq[dateSeq > as.Date("2000-01-02")]
dateSeq[weekdays(dateSeq) == "Monday"]
```

```{r}
identical(dateSeq[dateSeq > as.Date("2000-01-02")], dateSeq[which(dateSeq > as.Date("2000-01-02"))])
```

```{r}
library(microbenchmark)

microbenchmark(
  dateSeq[dateSeq > as.Date("2000-01-02")],
  times = 10000,
  unit = "us"
)

microbenchmark(
  dateSeq[which(dateSeq > as.Date("2000-01-02"))],
  times = 10000,
  unit = "us"
)
```

