---
title: "Exercises - Class 3: Vectors"
author: "Koen Plevoets"
date: "09-10-2023"
output: pdf_document
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(message = FALSE, warning = FALSE,
                      echo = FALSE)  # echo = TRUE for Solution
```

```{r initialize}
library(microbenchmark)
```

&nbsp;

### 1.
Create a 13x2 matrix `mat1` with the elements of the (built-in) vector `letters` in row-major order.

```{r rowMajor}
# Solution:
mat1 <- matrix(letters, nrow = 13, ncol = 2, byrow = TRUE)
mat1
```

### 2.
Create another 13x2 matrix `mat2` with the elements of `letters` in column-major order. How can you transform `mat2` in order to make it identical to `mat1` (from the previous exercise)?

```{r columnMajor, eval = knitr::opts_chunk$get("echo")}
# Solution:
mat2 <- matrix(letters, nrow = 13, ncol = 2)

identical(mat1, t(matrix(mat2, nrow = 2, ncol = 13)))
```

### 3.
Extract the 2nd column from `mat1` as a 13x1 matrix, not a vector. Do this for three different ways of indexing:

- positive integers
- negative integers
- logical vector

```{r drop, eval = knitr::opts_chunk$get("echo")}
# Solution:
mat1[, 2, drop = FALSE]
mat1[, -1, drop = FALSE]
mat1[, c(FALSE, TRUE), drop = FALSE]
```

### 4.
Compare the computation time of the three indexing operations from the previous exercise. You can use R's core function for timing but you will find the **microbenchmark** package more informative.

```{r time_drop, eval = knitr::opts_chunk$get("echo")}
# Solution:
system.time(mat1[, 2, drop = FALSE])
system.time(mat1[, -1, drop = FALSE])
system.time(mat1[, c(FALSE, TRUE), drop = FALSE])

microbenchmark(mat1[, 2, drop = FALSE])
microbenchmark(mat1[, -1, drop = FALSE])
microbenchmark(mat1[, c(FALSE, TRUE), drop = FALSE])
```

### 5.
Extract the five vowels from `mat1` using:

- a single index vector
- an index matrix

Again, time the computations of both operations.

```{r time_vowels, eval = knitr::opts_chunk$get("echo")}
# Solution:
system.time(mat1[c(1, 3, 5, 8, 11)])
system.time(mat1[rbind(c(1, 1), c(3, 1), c(5, 1), c(8, 1), c(11, 1))])

microbenchmark(mat1[c(1, 3, 5, 8, 11)],
               unit = "ns")
microbenchmark(mat1[rbind(c(1, 1), c(3, 1), c(5, 1), c(8, 1), c(11, 1))],
               unit = "ns")
# But:
microbenchmark(mat1[cbind(c(1, 3, 5, 8, 11), 1)],
               unit = "ns")
```

### 6.
What is the most efficient way to raise the first element of the following vector to the power `1`, the second element to the power `2` etc. and the last element to the power `10` (you can copy-paste the following code):

```{r show_vectorized, echo = TRUE}
vec1 <- 10:1
vec1
```

In other words, your result should be:

```{r vectorized}
# Solution:
vec1^(1:10)
```

### 7.
What is the most efficient way to raise the first element of the vector `vec1` (from the previous exercise) to the power `3`, the second element to the power `2`, the third element again to the power `3`, the fourth element again to the power `2` etc. up until the one-but-last element to the power `3` and the last element to the power `2`? In other words, your result should be:

```{r recycling}
# Solution:
vec1^c(3, 2)
```

### 8.
Consider the following matrix (you can copy-paste the code):

```{r show_recycling2, echo = TRUE}
mat3 <- matrix(seq(30, 270, by = 30), nrow = 3, ncol = 3)
mat3
```

How can you divide each row of `mat3` by the vector `c(2, 3, 5)` *without* using the the meta-function `sweep()`? Same question for dividing each column of `mat3` by `c(2, 3, 5)`, so your results should be:

```{r recycling2}
# Solution:
mat3 / c(2, 3, 5)
t(t(mat3) / c(2, 3, 5))
```

```{r comp_recycling2, eval = knitr::opts_chunk$get("echo")}
# Compare:
sweep(mat3, 1, c(2, 3, 5), "/")
sweep(mat3, 2, c(2, 3, 5), "/")
```

### 9.
Use a meta-function to generate all possible sums of the sides of two dice. The sides of one die can be represented as a vector (you can copy-paste the code):

```{r show_dice, echo = TRUE}
vec2 <- 1:6
vec2
```

Your result should be:

```{r dice}
# Solution:
outer(vec2, vec2, "+")
```

### 10.
The matrix from the previous exercise can be converted to a vector. There is another meta-function which allows you to compute the vector of all (36) sums of the sides of two dice. Compare the computation times of both commands.

```{r time_dice, eval = knitr::opts_chunk$get("echo")}
# Solution:
system.time(
  as.vector(outer(vec2, vec2, "+"))
)
system.time(
  apply(expand.grid(list(vec2, vec2)), 1, sum)
)

microbenchmark(
  as.vector(outer(vec2, vec2, "+"))
)
microbenchmark(
  apply(expand.grid(list(vec2, vec2)), 1, sum)
)
```

### 11.
Compare the computation time of R's core `cumsum()` function with `Reduce( . , accumulate = TRUE)`. Do this on this `vec3` (you can copy-paste the code):

```{r show_csum, echo = TRUE}
set.seed(123)
vec3 <- sample(vec1, size = 100, replace = TRUE)
head(vec3, n = 10)
tail(vec3, n = 10)
```

```{r time_csum, eval = knitr::opts_chunk$get("echo")}
# Solution:
system.time(
  cumsum(vec3)
  )
system.time(
  Reduce("+", vec3, accumulate = TRUE)
  )

microbenchmark(
  cumsum(vec3), 
  unit = "ns"
  )
microbenchmark(
  Reduce("+", vec3, accumulate = TRUE),
  unit = "ns"
  )
```

### 12.
Compare the computation times of three ways to sum every two consecutive elements in vector `vec3` (from the previous exercise):

- a `for` loop
- the `vapply()` meta-function
- another meta-function of the apply family, closely related to `vapply()`

In other words, the first element of your result `r vec3[1] + vec3[2]` should the sum of the first two elements `r vec3[1]` and `r vec3[2]` in `vec3`, the second element of the result `r vec3[2]  + vec3[3]` should be the sum of the second element `r vec3[2]` and third element `r vec3[3]`, the third element of the result `r vec3[3] + vec3[4]` needs to be the sum of the third element `r vec3[3]` and fourth element `r vec3[4]` in `vec3`, and so on:

```{r show_twos, echo = FALSE}
vapply(seq_len(length(vec3) - 1), function(i) sum(vec3[i:(i + 1)]), numeric(1))
```

This exercise assumes some basic familiarity with creating functions in R.

```{r time_twos, eval = knitr::opts_chunk$get("echo")}
# Solution:
system.time( {
  len <- length(vec3) - 1
  out <- numeric(len)
  for (i in seq_len(len)) {
    out[i] <- sum(vec3[i:(i + 1)])
  }
  out
  }
)
system.time(
 vapply(seq_len(length(vec3) - 1), function(i) sum(vec3[i:(i + 1)]), numeric(1)) 
)
system.time(
 sapply(seq_len(length(vec3) - 1), function(i) sum(vec3[i:(i + 1)])) 
)

microbenchmark( {
  len <- length(vec3) - 1
  out <- numeric(len)
  for (i in seq_len(len)) {
    out[i] <- sum(vec3[i:(i + 1)])
    }
  out
  },
  unit = "us")
microbenchmark(
 vapply(seq_len(length(vec3) - 1), function(i) sum(vec3[i:(i + 1)]), numeric(1)),
 unit = "us")
microbenchmark(
 sapply(seq_len(length(vec3) - 1), function(i) sum(vec3[i:(i + 1)])),
 unit = "us")
```
