---
title: "Exercises - Class 4: Functions"
author: "Koen Plevoets"
date: "11-10-2023"
output: pdf_document
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(message = FALSE, warning = FALSE,
                      echo = FALSE)  # echo = TRUE for Solution
```

```{r initialize}
library(methods)
library(microbenchmark)
```

&nbsp;

### 1.
[Harmonic numbers](https://en.wikipedia.org/wiki/Harmonic_number) are sums of the reciprocals of the positive integers. More specifically, the _n_'th harmonic number is:

$$
H_n = \sum_{i=1}^n \frac{1}{i}
$$
Create a function `Harmonic(n)` which computes the `n`th harmonic number. The body of this function can consist of a single line. An example is:

```{r Harmonic}
# Solution:
Harmonic <- function(n) sum(1 / (1:n))
```

```{r show_Harmonic, echo = TRUE}
Harmonic(5)
```

### 2.
Harmonic numbers are recursive:

$$
H_n = H_{n - 1} + \frac{1}{n}
$$

Write a recursive function `rHarmonic(n)` which computes the `n`th harmonic number recursively (in other words, the result should be the same as in the previous exercise).

```{r rHarmonic}
# Solution:
rHarmonic <- function(n) {
  if (n == 1) {
    1
  } else {
    rHarmonic(n - 1) + (1 / n)
  }
}
```

```{r show_rHarmonic, eval = knitr::opts_chunk$get("echo")}
rHarmonic(5)
```

### 3.
Compare the computation time of `Harmonic(n)` and `rHarmonic(n)` for the 1000th harmonic number (using R's core timing function).

```{r time_harmonic, eval = knitr::opts_chunk$get("echo"), error = TRUE}
# Solution:
n <- 1000
system.time(Harmonic(n))
system.time(rHarmonic(n))

# But:
n <- 1e6
system.time(Harmonic(n))
system.time(rHarmonic(n))
```

### 4.
Create a binary infix operator `%fill%` which replaces all `NA` values in the vector on the left-hand side by the value specified on the right-hand side. In other words, the function should work as follows:

```{r binary_fill}
# Solution:
`%fill%` <- function(x, y) {
  x[is.na(x)] <- y
  x
}
```

```{r binary_show, echo = TRUE}
vec1 <- c(5, 55, NA, -9, -99, NA, NA, -5, -9, 999)
vec2 <- vec1 %fill% 0
vec2
```

### 5.
Create a replacement function `fill()` which replaces all `NA` values in a vector by the value specified on the right-hand side. In other words, the function should work as follows:

```{r replace_fill}
# Solution:
`fill<-` <- function(x, value) {
  x[is.na(x)] <- value
  x
}
```

```{r replace_show, echo = TRUE}
vec1 <- c(5, 55, NA, -9, -99, NA, NA, -5, -9, 999)
fill(vec1) <- 0
vec1
```

### 6.
Compare the computation time of the `%fill%` operator and the `fill()` replacement function on the following large vector (you can copy-paste these two lines):

```{r show_specials, echo = TRUE}
set.seed(123)
vec3 <- sample(c(5, -9, NA), size = 1e6, replace = TRUE)
```

Do this using both R's core function and the package **microbenchmark** (replace the `NA`'s by `0`). 

```{r time_specials, eval = knitr::opts_chunk$get("echo"), cache = TRUE}
# Solution:
system.time(tmp <- vec3 %fill% 0)
system.time(fill(vec3) <- 0)

microbenchmark(tmp <- vec3 %fill% 0)
microbenchmark(fill(vec3) <- 0)
```

### 7.
Create two functions `full1()` and `full2()` which both extract the non-`NA` elements from a vector. One of both functions needs to make use of the functional `Filter()` (see last week). Compare their computations times on the `vec3` object from the previous exercises. You can use both R's core function and the **microbenchmark** package but for the latter you can specify a number of `times` <= 100.

**Note**: There is even a way to combine the `Filter()` function with the `Negate()` function. If you manage to find that one, then you can compare all three functions.

```{r defs_full}
# Solution, 1 / 2:
full1 <- function(x) {
  x[!is.na(x)]
}
full2 <- function(x) {
  Filter(function(y) !is.na(y), x)
}
full3 <- function(x) {
  Filter(Negate(is.na), x)
}
```

```{r demo_full, echo = TRUE}
vec1 <- c(5, 55, NA, -9, -99, NA, NA, -5, -9, 999)
full1(vec1)
full2(vec1)
```

```{r demo_neg3, eval = knitr::opts_chunk$get("echo")}
full3(vec1)
```

```{r time_full, eval = knitr::opts_chunk$get("echo"), cache = TRUE}
# Solution, 2 / 2:
system.time(full1(vec3))
system.time(full2(vec3))
system.time(full3(vec3))

microbenchmark(full1(vec3))
microbenchmark(full2(vec3))
microbenchmark(full3(vec3))
```

### 8.
One of the features of R's syntax is that a function name before the opening bracket `(` can actually be a variable. This allows uses such as the following:

```{r demo_varf, echo = TRUE}
f <- sum
f(vec2)
f <- mean
f(vec2)
f <- sd
f(vec2)

# Etc.
```

Create a function `twofun()` which has two formal arguments:

- `x` which can be a numeric vector
- `fun` which can be any function computing some descriptive statistic

The function `twofun()` should apply the function specified as `fun` to every two consecutive elements in the vector specified as `x`. In other words, if `fun = sum` then `twofun()` will compute the sum of every two consecutive elements in `x` like we have seen last week (see the last exercise of the previous week). However, `twofun()` allows for more flexibility:

```{r defs_varf}
# Solution:
twofun <- function(x, fun = sum) {
  vapply(seq_len(length(x) - 1),
         function(i) fun(x[i:(i + 1)]),
         numeric(1))
}
```

```{r demo_varfs, echo = TRUE}
twofun(vec1, fun = sum)
twofun(vec1, fun = mean)
twofun(vec1, fun = median)
twofun(vec1, fun = var)

# Etc.
```

Make sure that the function computes the sum by default:

```{r demo_varfs2, echo = TRUE}
twofun(vec1)
```

### 9.
R's functions for descriptive statistics have an argument `na.rm` and `var()` even has an argument `use`. Rewrite the `twosum()` function so that it can handle all of these extra arguments without any conflict.

```{r defs_ellipsis}
# Solution:
twofun2 <- function(x, fun = sum, ...) {
  vapply(seq_len(length(x) - 1),
         function(i) fun(x[i:(i + 1)], ...),
         numeric(1))
}
```

### 10.
User-defined functions in R have the type `closure` because they can be defined within another function (see the [Wikipedia page on closures](https://en.wikipedia.org/wiki/Closure_(computer_programming)) but a better description can be found in [chapter 10 of Hadley Wickham's _Advanced R_](https://adv-r.hadley.nz/function-factories.html)). Create a function `twofunfun()` which takes a statistical function as its single argument `fun` and which *returns another function* which applies `fun` to the two consecutive elements of any vector (i.e. like in the previous two exercises). In other words, `twofunfun()` should work as follows:

```{r defs_closure}
# Solution, 1 / 2:
twofunfun <- function(fun = sum) {
  out <- function(x, ...) {
    vapply(seq_len(length(x) - 1),
           function(i) fun(x[i:(i + 1)], ...),
           numeric(1))
  }
  out
}
```

```{r demo_closure, echo = TRUE}
twomean <- twofunfun(fun = mean)
twomean(vec1)
twovar <- twofunfun(fun = var)
twovar(vec1)

# Etc.
```

Because closures such as `twomean()` and/or `twovar()` are created within `twofunfun()`, they have the (temporary) execution environment of the latter as their environment:

```{r show_closure, echo = TRUE}
# Random memory location (so your environment name may be different):
environment(twomean)
# Random memory location (so your environment name may be different):
environment(twovar)
```

Such execution environments contain the values of the formal arguments (see slide 26) and that is why these environments should contain the value for the object `fun`. Show that this is indeed the case and get those values.

```{r local_variable, eval = knitr::opts_chunk$get("echo")}
# Solution, 2 / 2:
exists("fun", environment(twomean))
get("fun", environment(twomean))
exists("fun", environment(twovar))
get("fun", environment(twovar))
```

### 11.
Write an S3 method `freqs()` which counts how often every element appears in a vector. It should work for the S3 class `text` which is simply an extension of character vectors:

```{r init_S3, echo = TRUE}
vec4 <- c("a", "rose", "is", "a", "rose", "is", "a", "rose")
class(vec4) <- "text"
```

To be precise, the method `freqs()` should work as follows:

```{r defs_S3}
# Solution:
freqs <- function(x, ...) UseMethod("freqs")
freqs.text <- function(x) vapply(unique(x),
                                 function(j) sum(x == j),
                                 numeric(1))
```

```{r demo_S3, echo = TRUE}
freqs(vec4)
```

### 12.
Write a similar method `count()` for the S4 class `txt4` as well one as for the RC class `txtR` (we need to use different class and method names because otherwise R will mix them up; it is standard practice to choose just one OOP system and stick to it throughout your project).

```{r defs_Salt, results = "hide"}
# Solution:
setClass("txt4", contains = "character")
setGeneric("count", function(x) standardGeneric("count"))
setMethod("count", "txt4", function(x) {
  vapply(unique(x),
         function(j) sum(x == j),
         numeric(1))
  })

setRefClass("txtR", fields = c(x = "character"),
            methods = list(count = function() {
              vapply(unique(x),
                     function(j) sum(x == j),
                     numeric(1))
            }))
```

```{r show_Salt, echo = TRUE}
vec5 <- new("txt4", c("a", "rose", "is", "a", "rose", "is", "a", "rose"))
count(vec5)

vec6 <- new("txtR", x = c("a", "rose", "is", "a", "rose", "is", "a", "rose"))
vec6$count()
```

### 13.
Compare the computation time of the three methods with the computation time of simply using the `table()` function (which you might already have used in your methods but that is okay: in that case you will actually measure the computation time of the method dispatch). Do this comparison on a vector of a million randomly sampled letters from the (Roman) alphabet:

```{r show_data, echo = TRUE}
set.seed(123)
vec7 <- sample(letters, 1e6, replace = TRUE)

vec8 <- new("txt4", vec7)
vec9 <- new("txtR", x = vec7)
class(vec7) <- "text"
```

```{r time_data, eval = knitr::opts_chunk$get("echo"), cache = TRUE}
# Solution:
system.time(table(vec7, dnn = NULL))
system.time(freqs(vec7))
system.time(count(vec8))
system.time(vec9$count())

microbenchmark(table(vec7, dnn = NULL))
microbenchmark(freqs(vec7))
microbenchmark(count(vec8))
microbenchmark(vec9$count())
```

### 14.
A simple linear regression model estimates an **intercept** and a **slope** parameter of a regression line with which one variable can be explained/predicted in terms of another variable. For instance, the core R dataset `cars` has two variables _speed (mph)_ and _stopping distance (ft)_ and the second variable can be predicted/explained in terms of the first variable using the following intercept and slope parameters (estimated on the basis of 50 observations):

```{r init_scoping, echo = TRUE}
linmod <- lm(dist ~ speed, data = cars)
linmod
```

In other words, this linear regression model essentially states (using more digits for both parameters):

$$
dist = `r coef(linmod)[1]` + `r coef(linmod)[2]` \, \times \, speed
$$

This means that particular **predictions** for _stopping distance (ft)_ can be obtained by specifying certain values for _speed (mph)_ and plugging them into the equation of the linear regression model. Say, we would like the predicted values of _stopping distance (ft)_ for the _speed (mph)_ values _5_, _6_ and _7_. Modify the following two functions in order to obtain the proper result instead of the error message:

```{r demo_scoping, echo = TRUE, error = TRUE}
parameters <- function(mod) {
  coefs <- coefficients(mod)
}
predictions <- function(x, mod) {
  parameters(mod)
  coefs[1] + coefs[2] * x
}
predictions(x = 5:7, mod = linmod)
```

```{r scoping, eval = knitr::opts_chunk$get("echo")}
# Solution:
parameters <- function(mod) {
  coefficients(mod)    # No assignment is needed here
}
predictions <- function(x, mod) {
  coefs <- parameters(mod)    # R's scoping rules require assignment here
  coefs[1] + coefs[2] * x
}
predictions(x = 5:7, mod = linmod)
```
