---
title: "Exercises - Class 7: Text"
author: "Koen Plevoets"
date: "04-11-2023"
output: pdf_document
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(message = FALSE, warning = FALSE,
                      echo = FALSE)  # echo = TRUE for Solution
# This .Rmd file should be in the same folder alongside RStudio Project 'Text':
knitr::opts_knit$set(root.dir = ".//Text")
```

&nbsp;

The exercises can be carried out using the RStudio Project `Text`. It contains all necessary data files.

&nbsp;

## Exercises

### 1.
Consider the character vector `x`:

```{r show_split, echo = TRUE}
x <- c("Swan swam over the pond, Swim swan swim!",
    "Swam swam back again - Well swum swan!")
```

Split the strings in `x` into words, removing any *commas* or *hyphens*. Call the result `y`.

```{r split}
# Solution 1:
y <- strsplit(x, split = ",? -? ?")    # The argument split can be a regex!
y
```

### 2.
The .rda file `hafu` in the RStudio Project `Text` contains a data set `hafu` on manga characters. In the `Father` and `Mother` columns, some values have question marks after the country name, indicating that the author was uncertain about the nationality of the parent:

```{r show_hafu, echo = TRUE}
load("hafu.rda")
head(hafu[, -ncol(hafu)], n = 10)
```

Create two vectors `FatherUncertain` and `MotherUncertain` denoting whether or not there was a question mark in the `Father` or `Mother` column, respectively.

```{r hafu}
# Solution 2:
FatherUncertain <- with(hafu, grepl("\\?", Father))
MotherUncertain <- with(hafu, grepl("\\?", Mother))
```

### 3.
Create a new data frame `hafu2` by removing the question marks from the `Father` and `Mother` columns in `hafu` (i.e. replace them by an empty string).

```{r hafu2}
# Solution 3:
hafu2 <- within(hafu, {
  Father <- gsub("\\?", "", Father)
  Mother <- gsub("\\?", "", Mother)
})
```

&nbsp;

## Challenges

### 4.
The .rda file `Dataline` contains a vector `Dataline` that contains a single value. This value is the result of a webscraper. It contains 1500 fields, separated by a semicolon:

```{r show_Dataline, echo = TRUE}
load("Dataline.rda")
```

Create a vector `Fields` which contains the 1500 fields as separate values.

```{r Dataline}
# Solution 4:
Fields <- strsplit(Dataline, split = ";")[[1]]
```

### 5.
On the basis of the vector `Fields` create:

- A character vector `Label` containing the labels of every field: possible values are "Low", "Middle" and "High"
- A numeric vector `ID` containing the ID numbers of every field (i.e. the value following the underscore)
- A character vector `newLabel` containing labels constructed as `ID_xxx_Label`, with `xxx` the value of `ID` for that field and `Label` the value of `Label` for that field

```{r show_Fields, echo = TRUE}
head(Fields, n = 10)
```

```{r Fields}
# Solution 5:
Label <- gsub("ID.*", "", Fields)
ID <- as.numeric(gsub("[^_]+_", "", Fields))
newLabel <- paste("ID", ID, Label, sep = "_")
```

### 6.
The file genes.rda contains a named vector `genes`. This vector contains the function descriptions of a set of genes. The names of this vector give the COG codes of these genes:

```{r show_genes, echo = TRUE}
load("genes.rda")
head(genes, n = 10)
```

Use this vector to create a character vector `geneclass` with the values "Predicted", "Uncharacterized" and "Described". These are defined as follows:

- If the gene description contains "Uncharacterized", it is always called "Uncharacterized"
- If the gene description contains "Predicted" but not "Uncharacterized", it is called "Predicted"
- All other genes are called "Described"

```{r genes}
# Solution 6:
geneclass <- rep("Described", length(genes))
IDUnch <- grep("Uncharacterized", genes, ignore.case = TRUE)
IDPred <- grep("Predicted", genes, ignore.case = TRUE)
geneclass[IDPred] <- "Predicted"
geneclass[IDUnch] <- "Uncharacterized"
```

### 7.
Use the `genes` vector to find all COG codes for functions containing an enzyme name. An enzyme can be recognised by their name ending on "ase" or "ases". For this, you need to:

- Extract the COG code from the names: this consists of the numbers following the letters COG in the name
- Find at which positions in the vector genes an enzyme is described
- Use the information from the previous two steps to find only the COG codes of the enzymes

```{r enzymCOG}
# Solution 7:
COG <- gsub("\\D", "", names(genes))
isEnzyme <- grep("ases?\\b", genes)
enzymCOG <- COG[isEnzyme]
```

### 8.
The RStudio Project `Text` also contains two external files appletweets.txt and microsofttweets.txt. Each contains a selection of tweets scraped from Twitter/X and addressed to the helpdesk of either Apple or Microsoft.

Use the function `readLines()` to read in both files. Store the tweets for Apple in a vector `AppleTweets` and those for Microsoft in a vector `MSTweets`.

```{r readTweets}
# Solution 8:
AppleTweets <- readLines("appletweets.txt")
MSTweets <- readLines("microsofttweets.txt")
```

### 9.
Make sure that there are no empty character values (`""`) in either vector.

```{r removeEmpty}
# Solution 9:
AppleTweets <- AppleTweets[AppleTweets != ""]
MSTweets <- MSTweets[MSTweets != ""]
```

### 10.
Count the number of tweets that talk about iOS10. Think about possible spaces and capitalization.

```{r iOS10}
# Solution 10:
HASiOS10 <- sum(grepl("\\biOS\\s?10", AppleTweets, ignore.case = TRUE))
```

&nbsp;

## Extra exercises on dates and date-time objects

### 11.
Dates and Date-time objects in R are based on strings. Create the following objects (think about which type to use):

- A vector called `time1`: 2011-01-30
- A vector called `time2`: 30/01/11
- A vector called `time3`: Jan 30, 2011 14:45 UTC

You can you these strings as your starting point:

```{r show_times, echo = TRUE}
text1 <- "2011-01-30"
text2 <- "30/01/11"
text3 <- "Jan 30, 2011 14:45 UTC"
```

```{r times}
# Solution 11:
time1 <- as.Date(text1)
time2 <- as.Date(text2, format = "%d/%m/%y")
time3 <- as.POSIXct(text3, format = "%b %d, %Y %H:%M",
                    tz =  gsub(".+ (\\w{3}$)", "\\1", text3))
```

### 12.
Create a vector called `dateSeq` containing the dates:

- Starting from July, 1st 1989
- Ending at the last valid date before today
- With a periodicity of 3 months

```{r dateSeq}
# Solution 12:
dateSeq <- seq(as.Date("1989-07-01"),
               Sys.Date() - 1,
               by = "3 months")
```

### 13.
Find in `dateSeq`:

- All dates that are later than January 2nd, 2000
- All dates in dateSeq that fall on a Monday

```{r dateSeq2}
# Solution 13:
date212 <- dateSeq[which(dateSeq > as.Date("2000-01-02"))]
dateMon <- dateSeq[weekdays(dateSeq) == "Monday"]    # Depends on your Locale
```
