---
title: "Exercises - Class 6: Visualization"
author: "Koen Plevoets"
date: "29-10-2023"
output: pdf_document
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(message = FALSE, warning = FALSE,
                      echo = FALSE)  # echo = TRUE for Solution
```

```{r initialize}
library(lattice)
library(ggplot2)
```

&nbsp;

### 1.
R has a core function `coplot()` to do multi-panel conditioning. Use it on the (core) dataset `mtcars` to create the following plot:

```{r coplot}
coplot(mpg ~ disp | factor(cyl), data = mtcars, show.given = FALSE)
```

### 2.
Add a LOESS smoother to the previous conditioning plot:

```{r smooth}
coplot(mpg ~ disp | factor(cyl), data = mtcars, show.given = FALSE,
       panel = panel.smooth)
```

### 3.
Use the `par()` function to check R's different coordinate systems: verify that changing one graphical parameter indeed affects other graphical parameters.

```{r pars, eval = knitr::opts_chunk$get("echo")}
# The outer margins are by default set to zero in all coordinate systems:
par("oma")    # Lines of text (bottom, left, top, right)
par("omi")    # Inches (bottom, left, top, right)
par("omd")    # Proportion of device region (left, right, bottom, top)
# This means that the figure region is set to its maximal size:
par("fin")    # Inches (width, height)
par("fig")    # Proportion of device region (left, right, bottom, top)
# The figure region is the sum of the figure margins and the plot region:
par("mai")    # Inches (bottom, left, top, right)
par("pin")    # Inches (width, height)
# Ergo, par("fin") ==
par("pin") + c(sum(par("mai")[c(2, 4)]), sum(par("mai")[c(1, 3)]))
# A corresponding coordinate system for par("mai") is:
par("mar")    # Lines of text (bottom, left, top, zero)
# A corresponding coordinate system for par("pin") is:
par("plt")    # Proportion of figure region (left, right, bottom, top)
# Changing e.g. the outer margins in one coordinate system changes the
# corresponding coordinate systems:
par(oma = c(0.5, 0.5, 0.5, 0.5))
par("omi")
par("omd")
# However, it can also change a graphical parameter for another component:
par("fin")
# But not:
par("fig")
# Because the figure margins are not affected by changing the outer margins...
par("mar")
par("mai")
# ...the plot region must change:
par("pin")
par("plt")
# Similarly, changing the figure margins while keeping the figure region constant
# will change the plot region and vice versa.
# Etc.
# Undo the change:
par(oma = c(0, 0, 0, 0))
```

### 4.
Use the **lattice** package to create a simple scatter plot of all the data for the `mpg` and `disp` variable of the `mtcars` dataset. Store the plot in a graphical objects called `plt1`. Compute the byte size of that graphical object.

```{r xyplot}
plt1 <- xyplot(mpg ~ disp, data = mtcars)
plt1
```

```{r size_xyplot, eval = knitr::opts_chunk$get("echo")}
object.size(plt1)
```

### 5.
Use the **lattice** package to create the same conditioning plot as in Exercise 1. Store it in a graphical object `plt2` and compute the byte size of this object:

```{r xypanel}
plt2 <- xyplot(mpg ~ disp | factor(cyl), data = mtcars, layout = c(2, 2))
plt2
```

```{r size_xypanel, eval = knitr::opts_chunk$get("echo")}
object.size(plt2)
```

### 6.
Add a LOESS smoother to the previous **lattice** plot (in other words, create a similar plot as in Exercise 2). Store it in a graphical object `plt3` and compute the byte size of this object:

```{r xysmooth, warning = FALSE}
plt3 <- xyplot(mpg ~ disp | factor(cyl), data = mtcars, layout = c(2, 2),
               panel = function(x, y, ...) {
                 panel.xyplot(x, y, ...)
                 panel.loess(x, y, ...)
               })
# There can be some warnings about the LOESS smoother...
plt3
```

```{r size_xysmooth, eval = knitr::opts_chunk$get("echo")}
object.size(plt3)
```

### 7.
Use the **ggplot2** package to create a graphical object `plt4` in which the `mpg` and `disp` variables of the `mtcars` dataset are mapped onto the `y` and `x` aesthetic, respectively. Compute the byte size of that graphical object.

```{r ggplot2_object}
plt4 <- ggplot(data = mtcars, mapping = aes(x = disp, y = mpg))
```

```{r size_object, eval = knitr::opts_chunk$get("echo")}
object.size(plt4)
```

### 8.
Add a point geom to the graphical object, so you get a similar scatter plot as in Exercise 4. Store this in a graphical object `plt5` and compute its byte size.

```{r ggplot2_geom}
plt5 <- plt4 + geom_point()
plt5
```

```{r size_geom, eval = knitr::opts_chunk$get("echo")}
object.size(plt5)
```

### 9.
Add a facet layer for the `cyl` variable to the previous graphical object. Store it in an object `plt6` and compute the byte size of this object.

```{r ggplot2_facet}
plt6 <- plt5 + facet_wrap(facets = ~ factor(cyl), nrow = 2, as.table = FALSE)
plt6
```

```{r size_facet, eval = knitr::opts_chunk$get("echo")}
object.size(plt6)    # No difference !
```

### 10.
Add a layer with a LOESS smoother to the graphical object of the previous exercise (this may mean that the geom layer is automatically adjusted). Store this into a graphical object `plt7` and compute its byte size.

```{r ggplot2_stat, warning = FALSE}
plt7 <- plt6 + stat_smooth(method = "loess", se = FALSE)
# There can be some warnings about the LOESS smoother...
plt7
```

```{r size_stat, eval = knitr::opts_chunk$get("echo")}
object.size(plt7)
```
