---
title: "Edgar Anderson's Iris Data"
output:
    pdf_document:
        toc: yes
---

```{r settings, include = FALSE}
options(width = 100)
```

## Description

This famous ([Fisher's](https://en.wikipedia.org/wiki/Ronald_Fisher) or
[Anderson's](https://en.wikipedia.org/wiki/Edgar_Anderson)) iris data set gives
the measurements in centimeters of the variables sepal length and width and
petal length and width, respectively, for 50 flowers from each of 3 species of
iris. The species are _Iris setosa_, _versicolor_, and _virginica_.

## Usage

`iris`

## Format

`iris` is a data frame with 150 cases (rows) and 5 variables (columns) named:

- `Sepal.Length`
- `Sepal.Width`
- `Petal.Length`
- `Petal.Width`
- `Species`

## Source

Anderson, Edgar (1935). "The irises of the Gaspe Peninsula."
_Bulletin of the American Iris Society_, __59__: 2--5.  
Fisher, Ronald A. (1936). "The use of multiple measurements in taxonomic problems."
_Annals of Eugenics_, __7__ (Part II): 179--188.  

## Examples

We investigate the Sepal and Petal leaves for the three species in the Iris data:

```{r overview}
summary(iris)
```

To examine the Sepal leaves, we select the length and the width:

```{r Sepal_vars}
llen <- iris$Sepal.Length
lwid <- iris$Sepal.Width
```

Then we plot the data:

```{r Sepal_plot}
plot(llen, lwid, xlab = "Length", ylab = "Width", 
     pch = 20, col = as.numeric(iris$Species))
legend("bottomright", legend = levels(iris$Species), col = 1:3, pch = 20)
```

We can also select the Petal leaves:

```{r Petal_vars}
llen <- iris$Petal.Length
lwid <- iris$Petal.Width
```

This gives us the following plot:

```{r Petal_plot, ref.label = "Sepal_plot", echo = FALSE}

```
